Comment: Inferring Trade Costs from Trade Booms and Trade Busts

### 2017-09-15 Paper/Manuscript.tex
Pour tableau 1 et Our method
Comme dans le papier soumis

\documentclass{article}
\usepackage{authblk}
\usepackage{natbib}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[pdftex]{graphicx}  
%Langue utilisée
\usepackage[english]{babel}
\usepackage[T1]{fontenc}
\emergencystretch=8em
\usepackage[margin=1in]{geometry}

@@

\author[1]{Guillaume Corlay}
\author[2]{Stéphane Dupraz}
\author[3]{Claire Labonne}
\author[1]{Anne Muller}
\author[4]{Céline Antonin}
\author[5]{Guillaume Daudin\footnote{Corresponding author: guillaume.daudin@dauphine.fr. The authors thank Marc Adam, Marcelo Olarreaga and anonymous referees for their comments and questions}}
\affil[1] {ENSAE, French National School of Statistics and Economic Administration, 3, avenue Pierre Larousse, 92245 Malakoff, FRANCE}
\affil[2] {Columbia University, 10025 New York NY, USA}
\affil[3] { Paris School of Economics / Université Paris 1 Panthéon Sorbonne -- ACPR - Banque de France, 75009, PARIS, FRANCE}
\affil[4] {Sciences Po, Observatoire Français des Conjonctures Économiques (OFCE), 75007, PARIS, FRANCE}
\affil[5] {Université Paris-Dauphine, PSL Research University, LEDa, 75016 PARIS, FRANCE

@@

\begin{abstract}
\cite{JMN2011} offer a method to measure trade costs that relies exclusively on bilateral exports and GDP statistics.
They argue that the reduction in trade costs was the main driving force of trade growth during the first globalization (1870-1913), whereas economic expansion was the main driving force during the second globalization (1950-2000).
This potentially important result is driven by the use of an \textit{ad hoc} aggregation method of bilateral trade costs to the country and global level.
What \cite{JMN2011} capture is the difference in the relative starting trade of dyads experiencing faster trade growth in the first and second globalization.
More generally, we cast doubts on the possibility to reach conclusions of such nature with a method that excludes a priori changes in non-trade costs determinants of openness rates and hence can only rephrases the information contained in openness ratios.
\end{abstract}

@@

Still, when we decompose the level of exportations of a country simply as the product of its GDP and its openness ratio (see table \ref{OR}) we find that the increase in exportations is mainly explained by GDP growth for the large majority of the 27 countries in the data.\footnote{We follow the authors in interpreting log differences as percentages. However, one should keep in mind that given the size of the changes, this is a very inexact approximation.}
On average, this decomposition attributes 74\% (183/246) of the growth in trade to the increase in GDP in the second globalization and 60\% in the first globalization.
This is very much in contrast to  Jacks \textit{et al.}\footnote{Contributions of growth in income similarity and of change in multilateral factors are negative.} :

@@

\begin{table}
\centering
\begin{tabular}{|r|r|r|r|r|r|r|}
\hline
 & \multicolumn{3}{|c|}{First Globalization 1870-1913} & \multicolumn{3}{|c|}{Second Globalization 1950-2000}\\
\hline
 & Exports & GDP & Openness Ratio & Exports & GDP & Openness Ratio\\
\hline
Argentina & 292 & 251 & 40 & 116 & 132 & -16\\
\hline
Austria & 89 & 102 & -13 & 379 & 185 & 194\\
\hline
Australia & 174 & 152 & 22 & 114 & 191 & -77\\
\hline
Belgium & 195 & 86 & 109 & 251 & 150 & 101\\
\hline
Brazil & 174 & 101 & 73 & 184 & 239 & -55\\
\hline
Canada & 233 & 170 & 64 & 258 & 192 & 66\\
\hline
Denmark & 200 & 113 & 87 & 230 & 142 & 88\\
\hline
France & 117 & 70 & 48 & 260 & 172 & 87\\
\hline
Germany & 168 & 119 & 48 & 391 & 175 & 216\\
\hline
Greece & 140 & 99 & 42 & 227 & 218 & 9\\
\hline
India & 147 & 41 & 105 & 161 & 216 & -54\\
\hline
Indonesia & 220 & 87 & 133 & 263 & 232 & 31\\
\hline
Italy & 151 & 83 & 69 & 335 & 188 & 147\\
\hline
Japan & 337 & 104 & 233 & 448 & 281 & 167\\
\hline
Mexico & 189 & 143 & 46 & 296 & 238 & 59\\
\hline
Netherlands & 230 & 92 & 138 & 298 & 173 & 125\\
\hline
New Zealand & 184 & 186 & -2 & 307 & 133 & 174\\
\hline
Norway & . & . & . & 70 & 181 & -111\\
\hline
Philippines & 218 & 92 & 126 & 237 & 213 & 24\\
\hline
Portugal & 74 & 57 & 17 & 319 & 208 & 111\\
\hline
Spain & 171 & 76 & 96 & 496 & 230 & 266\\
\hline
Sri Lanka & 172 & 92 & 81 & 39 & 201 & -161\\
\hline
Sweden & 150 & 92 & 58 & 240 & 134 & 106\\
\hline
Switzerland & 105 & 108 & -3 & 250 & 132 & 118\\
\hline
UK & 125 & 81 & 44 & 196 & 122 & 74\\
\hline
Uruguay & 261 & 165 & 96 & 48 & 94 & -46\\
\hline
USA & 208 & 166 & 42 & 241 & 170 & 72\\
\hline
 &  &  &  &  &  & \\
\hline
Average & \textbf{178} & \textbf{107} & \textbf{71} & \textbf{246} & \textbf{183} & \textbf{63}\\
\hline
\end{tabular}
\caption{Decomposition of the growth of exportations between
GDP growth and Openness ratio growth, log differences (interpreted as percentages).
Figures for Norway are not given because the dissolution of the
union between Norway and Sweden (1905) makes them
meaningless.}\label{OR}
\end{table}

@@

\cite{AW2003}'s model can be used to calculate the aggregate
trade costs a single country faces in its trade with all
its trading partners. Indeed, using a two-countries model where
all trade partners of country $i$ are treated as one single
country provides an aggregate measure of trade costs that does
not rely on any \textit{ad hoc} aggregation method but instead on
microfoundations.

@@

where $x_{ip}$ and $x_{pi}$ are respectively real exports and
imports of $i$ from and to its trade partners in the data, and
$x_{pp}$ is the volume of trade within and between the trading
partners present in the data. This `domestic trade' variable
now
includes cross-border trade. The problem with this method is that it is very data-intensive. It requires the knowledge of all bilateral trade flows between the countries in the data. \\

Compare this with Jacks \textit{et al.}'s  method : 
\begin{eqnarray}
1+\tau_i&=& \sum_{j \neq i}  {\frac{y_i}{\sum_{j \neq i} y_i} \left(1+\tau_{ij}\right)}
= \sum_{j \neq i}  {\frac{y_i}{\sum_{j \neq i} y_i} \left( \frac{x_{ii} x_{jj}}{x_{ij} x_{ji}}\right)^{\frac{1}{2(\sigma-1)}} }
\end{eqnarray}

Using this measure of trade costs, we provide the same
decomposition in four terms of the increase in trade flows
during the two globalizations as in the authors' article for
the countries with the most trade partners available in the
data (France (24), the UK (25), the USA (23)). We also provide
the unweighted and end-of-period-GDP-weighted averages for all
countries in the sample. Results are displayed in tables
\ref{glob1} and \ref{glob2} along with the results with the
Jacks \textit{et al.}'s method.\footnote{They include unweighted and
end-of-period-GPD-weighted averages over country dyads of
bilateral trade costs, such as provided in \cite{JMN2011}. We
also display for France, the UK and the USA the trade cost
measure averaged over trading partners faced by one country, as
the results are presented (with a smaller data set) in
\cite{JMN2008, JMN2010}. We also provide both averages of this
measure over all countries in the data set.}\\

>>>>

\documentclass{article}
\usepackage{authblk}
\usepackage{natbib}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[pdftex]{graphicx}  
%Langue utilisée
\usepackage[english]{babel}
\usepackage[T1]{fontenc}
\emergencystretch=8em
\usepackage[margin=1in]{geometry}
\usepackage{booktabs}
\usepackage{tabularx}

@@

\author[1]{Guillaume Corlay}
\author[2]{Stéphane Dupraz}
\author[3]{Claire Labonne}
\author[1]{Anne Muller}
\author[4]{Céline Antonin}
\author[5]{Guillaume Daudin\footnote{Corresponding author: guillaume.daudin@dauphine.fr. The authors thank Marc Adam, Marcelo Olarreaga and anonymous referees for their comments and questions. Marc Adam pointed to us a slight data incoherence in Jacks \textit{et al.}'s .dta files that we have corrected in the following analysis.}}
\affil[1] {ENSAE, French National School of Statistics and Economic Administration, 3, avenue Pierre Larousse, 92245 Malakoff, FRANCE}
\affil[2] {Columbia University, 10025 New York NY, USA}
\affil[3] { Paris School of Economics / Université Paris 1 Panthéon Sorbonne -- ACPR - Banque de France, 75009, PARIS, FRANCE}
\affil[4] {Sciences Po, Observatoire Français des Conjonctures Économiques (OFCE), 75007, PARIS, FRANCE}
\affil[5] {Université Paris-Dauphine, PSL Research University, LEDa, 75016 PARIS, FRANCE

@@

\begin{abstract}
\cite{JMN2011} offer a method to measure trade costs that relies exclusively on bilateral exports and GDP statistics.
They argue that the reduction in trade costs was the main driving force of trade growth during the first globalization (1870-1913), whereas economic expansion was the main driving force during the second globalization (1950-2000).
This potentially important result is driven by the use of an \textit{ad hoc} aggregation method of bilateral trade costs to the country and global level.
What \cite{JMN2011} capture is the difference in the relative starting trade of dyads experiencing faster trade growth in the first and second globalization.
More generally, we cast doubts on the possibility to reach conclusions on aggregate costs with a method that excludes a priori changes in non-trade costs determinants of openness rates and hence can only rephrases the information contained in them.
\end{abstract}

@@

Still, when we decompose the level of exportations of a country simply as the product of its GDP and its openness ratio (see table \ref{OR}) we find that the increase in exportations is mainly explained by GDP growth for the large majority of the 27 countries in the data.\footnote{We follow the authors in interpreting log differences as percentages. However, one should keep in mind that given the size of the changes, this is a very inexact approximation.}
On average, this decomposition attributes 74\% (183/246) of the growth in trade to the increase in GDP in the second globalization and 62\% in the first globalization.
This is very much in contrast to  Jacks \textit{et al.}\footnote{Contributions of growth in income similarity and of change in multilateral factors are negative.} :

@@

\begin{table}[tbp] \centering
\newcolumntype{C}{>{\centering\arraybackslash}X}

\begin{tabularx}{\textwidth}{lCCCCCC}

\toprule
& \multicolumn{3}{c}{First Globalization 1870-1913} & \multicolumn{3}{c}{Second Globalization 1950-2000}\\
{}&{Exports}&{GDP}&{Openness Ratio}&{Exports}&{GDP}&{Openness Ratio} \tabularnewline
\midrule\addlinespace[1.5ex]
ARG&292&251&40&116&132&-16 \tabularnewline
AUH&89&102&-13&379&185&194 \tabularnewline
AUS&174&152&22&114&191&-77 \tabularnewline
BEL&195&86&109&251&150&101 \tabularnewline
BRA&174&101&73&184&239&-55 \tabularnewline
CAN&233&170&64&258&192&66 \tabularnewline
DEN&200&113&87&230&142&88 \tabularnewline
FRA&117&70&48&260&172&87 \tabularnewline
GER&168&119&48&391&175&216 \tabularnewline
GRE&140&99&42&227&218&9 \tabularnewline
IND&147&41&105&161&216&-54 \tabularnewline
INN&220&87&133&263&232&31 \tabularnewline
ITA&151&83&69&335&188&147 \tabularnewline
JAP&337&104&233&448&281&167 \tabularnewline
MEX&189&143&46&296&238&59 \tabularnewline
NET&230&92&138&298&173&125 \tabularnewline
NEW&184&186&-2&307&133&174 \tabularnewline
NOR&&&&70&181&-111 \tabularnewline
PHI&218&92&126&237&213&24 \tabularnewline
POR&74&57&17&319&208&111 \tabularnewline
SPA&171&76&96&496&230&266 \tabularnewline
SRI&172&92&81&39&201&-161 \tabularnewline
SWE&150&92&58&240&134&106 \tabularnewline
SWI&105&108&-3&250&132&118 \tabularnewline
UK&125&81&44&196&122&74 \tabularnewline
URU&261&165&96&48&94&-46 \tabularnewline
USA&208&166&42&241&170&72 \tabularnewline
\textbf{Average}&\textbf{182}&\textbf{113}&\textbf{69}&\textbf{246}&\textbf{183}&\textbf{63} \tabularnewline
\bottomrule \addlinespace[1.5ex]

\end{tabularx}
\caption{Decomposition of the growth of exportations between
GDP growth and Openness ratio growth, log differences (interpreted as percentages).
Figures for Norway are not given because the dissolution of the
union between Norway and Sweden (1905) makes them
meaningless.}\label{OR}
\end{table}

@@

\cite{AW2003}'s model can be used to calculate the aggregate
trade costs a single country faces in its trade with all
its trading partners. Indeed, using a two-countries model where
all trade partners of country $i$ are treated as one single
country provides an aggregate measure of trade costs that does
not rely on any \textit{ad hoc} aggregation method but instead on
microfoundations. The drawback of this method is that it is very data-intensive. It requires the knowledge of all bilateral trade flows between the countries in the data.

@@

where $x_{ip}$ and $x_{pi}$ are respectively real exports and
imports of $i$ from and to its trade partners in the data, and
$x_{pp}$ is the volume of trade within and between the trading
partners present in the data. This `domestic trade' variable
now
includes cross-border trade. \\

Compare this with Jacks \textit{et al.}'s  method to measure the mean trade cost faced by a country : 
\begin{eqnarray}
1+\tau_i&=& \sum_{j \neq i}  {\frac{y_i}{\sum_{j \neq i} y_i} \left(1+\tau_{ij}\right)}
= \sum_{j \neq i}  {\frac{y_i}{\sum_{j \neq i} y_i} \left( \frac{x_{ii} x_{jj}}{x_{ij} x_{ji}}\right)^{\frac{1}{2(\sigma-1)}} }
\end{eqnarray}

To illustrate the practical difference between these two methods, we provide the same decomposition in four terms of the increase in trade flows during the two globalizations as in the authors' article for the countries with the most trade partners available in the data (France (24), the UK (25), the USA (23)).
This requires studying only trade between Belgium, Canada, France, Germany, India, Italy, Spain, United Kingdom and United States of America, as this is the largest set in the data where all bilateral trade flows are available. 
We also provide the unweighted and end-of-period-GDP-weighted averages for all countries in the sample. 
Results are displayed in tables \ref{glob1} and \ref{glob2} along with the results with the
Jacks \textit{et al.}'s method.\footnote{They include unweighted and
end-of-period-GPD-weighted averages over country dyads of
bilateral trade costs, such as provided in \cite{JMN2011}.
We also display for France, the UK and the USA the trade cost measure averaged over trading partners faced by one country, as the results are presented (with a smaller data set) in
\cite{JMN2008, JMN2010}.
We also provide both averages of this measure over all countries in the data set.}\\
